A free/open-source Kazakh-Tatar machine translation system

### 2013-04-22 paper/paper.tex
aaaand centered
minor stuff
+1 lexical selection rule
another lrx rule

\begin{figure}
\begin{lstlisting}
<rule> 
  <match lemma="ақпарат" tags="n.attr"/> 
  <match lemma="құрал" tags="n.*.px3sp.*">
    <select lemma="чара"/>
  </match> 
</rule>
\end{lstlisting}

\todo{A couple more lexical-selection rules }
\caption{A lexical selection rule that selects чара as the translation of құрал if part of a compound with ақпарат.}
\label{fig:lrx}
%\vspace{-1em}
\end{figure}

@@

\begin{figure}[htbp]
	\begin{tabular}{lll}
		\toprule
		form & current & expected \\\midrule
		ди\tag{<v>}\tag{<tv>}\tag{<pres>}\tag{<p3>}\tag{<sg>} & дия & ди \\
		йөр\tag{<v>}\tag{<iv>}\tag{<pres>}\tag{<p3>}\tag{<sg>} & йөрә & йөри \\
		кияү\tag{<n>}\tag{<px3sp>}\tag{<nom>} & кияүе & кияве \\
		укы\tag{<v>}\tag{<iv>}\tag{<ger>}\tag{<nom>} & укыу & уку \\\bottomrule
	\end{tabular}
	\caption{Examples of some phonological problems in the Tatar transducer.}
	\label{tab:phonerrors}
\end{figure}

@@

\begin{table}
  \begin{center}
  \begin{tabular}{ccrrr}
  \toprule
   Corpus                 & Direction           & Tokens  & OOV & WER (\%) \\
  \midrule
  \texttt{devel} & kaz$\rightarrow$tat & 2457    & 2       & 15.19 \\
  \bottomrule
  \texttt{test} & kaz$\rightarrow$tat & 2862    & 43      & 36.57 \\
  \end{tabular}
    \caption{Word error rate over two corpora. The \texttt{devel} `development' corpus was used during development 
       to identify possible disambiguation, lexical-selection and transfer rules. The \texttt{test} corpus
       was not seen during development. The OOV column gives the number of out-of-vocabulary (unknown) words.}
 %\vspace{-2.5em}
    \label{table:wer-development}
  \end{center}
\end{table}

>>>>

\begin{figure}
\begin{lstlisting}
<rule> 
  <match lemma="ақпарат" tags="n.attr"/> 
  <match lemma="құрал" tags="n.*.px3sp.*">
    <select lemma="чара"/>
  </match> 
</rule>
\end{lstlisting}
\begin{lstlisting}
  <rule>
<!-- 61. Германия құрамасы Еуропа чемпионатын үш рет ұтып алған.
     61. Германия җыелмасы Аурупа чемпионатын өч тапкыр откан. -->
    <match tags="num"/>
    <match lemma="рет" tags="n.*"><select lemma="тапкыр"/></match>
  </rule>
\end{lstlisting}
\begin{lstlisting}
  <rule>
    <match lemma="рухани" tags="adj"/>
    <match lemma="өмір" tags="n.*"><select lemma="тормыш"/></match>
  </rule>
\end{lstlisting}
\todo{A couple more lexical-selection rules }
\caption{A lexical selection rule that selects чара as the translation of құрал if part of a compound with ақпарат.}
\label{fig:lrx}
%\vspace{-1em}
\end{figure}

@@

\begin{figure}[htbp]
	\centering
	\begin{tabular}{lll}
		\toprule
		form & current & expected \\\midrule
		ди\tag{<v>}\tag{<tv>}\tag{<pres>}\tag{<p3>}\tag{<sg>} & дия & ди \\
		йөр\tag{<v>}\tag{<iv>}\tag{<pres>}\tag{<p3>}\tag{<sg>} & йөрә & йөри \\
		кияү\tag{<n>}\tag{<px3sp>}\tag{<nom>} & кияүе & кияве \\
		укы\tag{<v>}\tag{<iv>}\tag{<ger>}\tag{<nom>} & укыу & уку \\\bottomrule
	\end{tabular}
	\caption{Examples of some phonological problems in the Tatar transducer.}
	\label{tab:phonerrors}
\end{figure}

@@

\begin{table}
  \begin{center}
  \begin{tabular}{ccrrr}
  \toprule
   Corpus                 & Direction           & Tokens  & OOV & WER (\%) \\
  \midrule
  \texttt{devel} & kaz$\rightarrow$tat & 2457    & 2       & 15.19 \\
  \texttt{test} & kaz$\rightarrow$tat & 2862    & 43      & 36.57 \\
  \bottomrule
  \end{tabular}
    \caption{Word error rate over two corpora. The \texttt{devel} `development' corpus was used during development 
       to identify possible disambiguation, lexical-selection and transfer rules. The \texttt{test} corpus
       was not seen during development. The OOV column gives the number of out-of-vocabulary (unknown) words.}
 %\vspace{-2.5em}
    \label{table:wer-development}
  \end{center}
\end{table}
